---
title: "ASR Replication Preliminary - Felony"
author: "Ryan Larson - UMN"
date: "`r format(Sys.time(), '%m/%d/%Y')`"
output: pdf_document
fontsize: 11pt
editor_options: 
  chunk_output_type: console
---

This analysis represents a preliminary attempt to replicate and extend the Harris et. al. ASR paper. The Minnesota data is situated at the case level from 2010-2015.  The full data series extends back to 2004, but VIBES data in the two most populous MN counties, Hennepin and Ramsey, are missing and inconsistent until 2010. Additionally, while data at the conviction level would be more comparable, we do not have data at the conviction level for much of the VIBES data. This analysis uses the felony only data which more closely reflects the ASR paper's use of Washington Superior Court data. We also include an extention set of mixed models to Native American populations, with hypothesized interactions with alcohol-related and hunting and fisheries violations to examine if the courtesty stigma and steretype congruence effects found in the Black and Latino models extend to Native American populations.  


```{r, warning=F, include=F, message=F}
##########################################
# Multi-State Study of Monetary Sanctions
# Ryan Larson, RA, UMN
##########################################

#TO DO
#model with all interactions
#add in other covariates(e.g., sentence), make different models
#make more complex crim history var
#L2 poverty - will make from CPS?
#dynamic level-2 covariates?
#county fixed effects - make sure level-2 is time varying


#packages
library(readr)
library(dplyr)
library(ggplot2)
library(lme4)
library(lmerTest)
library(stargazer)
library(car)

#load data - constructed from MCAO .txt files
monsanc <- read_csv("~/MonSanc/monsanc.csv") 

#recoding and filtering
monsanc.short <- monsanc %>%
  mutate(total_ff = total_ff*adj, #adjust dollars to Jan. 2018
         total_ff_stand = scale(total_ff),
         total_ff_log = log((total_ff+1)*adj), #log+1 DV
         min_perc = (min_pop/total_pop)*100,
         black_perc = (black_pop/total_pop)*100,
         hisp_perc = (hisp_pop/total_pop)*100, 
         na_perc = (na_pop/total_pop)*100) %>%
  filter(file_year >= 2010) #restricting to cases started from 2010-2015

#make factors into binary indicators to match analyses/tables from paper
monsanc.short <- monsanc.short %>% 
  mutate(charge_degree = ifelse(charge_degree=="converted"|charge_degree=="other", NA, charge_degree),
         felony = ifelse(charge_degree=="felony", 1, 0), 
         gross.mis = ifelse(charge_degree=="gross misdemeanor", 1, 0),
         mis = ifelse(charge_degree=="misdemeanor", 1, 0),
         petty.mis = ifelse(charge_degree=="petty misdemeanor", 1, 0),
         white = ifelse(race.squish=="white", 1, 0),
         asian = ifelse(race.squish=="asian", 1, 0),
         black = ifelse(race.squish=="black", 1, 0),
         hispanic = ifelse(race.squish=="hispanic", 1, 0),
         nativeam = ifelse(race.squish=="nat. am.", 1, 0),
         other.race = ifelse(race.squish=="other", 1, 0),
         violent = ifelse(charge_offense=="violent", 1, 0),
         alcohol.dui = ifelse(charge_offense=="alcohol/dui", 1, 0),
         drug = ifelse(charge_offense=="drug", 1, 0),
         hunt.fish = ifelse(charge_offense=="hunt/fish", 1, 0),
         other.offense = ifelse(charge_offense=="other", 1, 0),
         male = ifelse(gender=="M", 1,0)) %>%
  filter(age >= 15) #filtering cases with likely error ages (n=620)

#felony filter
felony <- monsanc.short %>% filter(charge_degree=="felony")
```

## Distribution of Fines and Fees

```{r, message=F, warning=F, echo=F}
fel.mean <- mean(felony$total_ff, na.rm=T)

ggplot(subset(felony, total_ff <=2000), aes(x=total_ff)) + 
  geom_histogram(colour="black", fill="white") +
  labs(title = "Distribution of Fine and Fee Amount < $2000", x = "Total Fine/Fee Order ($)", y = "Frequency",
       caption = "Note: Adjusted to Jan. 2018 dollars")+
  theme_minimal()+
  geom_vline(aes(xintercept = fel.mean), color="blue", linetype="dashed")

```


\pagebreak

## Descriptive Statistics

```{r,echo=F, results='asis'}
stargazer(as.data.frame(felony[,c(23,53,55,116,105,106,107,108,109,110,52,111,112,113,114,115,97,80,94)]), 
          type="latex", style="asr", title="Descriptive Statistics - Felony Data",
          summary=T, covariate.labels=c("Total Fine/Fee", "Priors", "Age", 
                                                  "Male", "White","Asian", "Black", 
                                                  "Hispanic", "Native Am.", "Other Race", "Trial", 
                                                  "Violent","Alcohol/DUI", "Drug", "Hunt/Fish", 
                                                  "Other Offense", "Percent Minority", 
                                                  "Percent Vote Republican", "Percent Law and Justice"), header = F)
```

\pagebreak

## Bivariate Correlations

```{r, include=F}
corr<-round(cor(felony[,c(96,53,55,116,105,106,107,108,109,110,52,111,112,113,114,115,97,80,94)], use="pairwise.complete.obs", method="pearson" ),2)
corr[upper.tri(corr)] <-""
upper <- as.matrix(corr)
rownames(upper) <- c("Total Fine/Fee (log)", "Priors", "Age", 
                                                  "Male", "White","Asian", "Black", 
                                                  "Hispanic", "Native Am.", "Other Race", "Trial", 
                                                  "Violent","Alcohol.DUI", "Drug", "Hunt.Fish", 
                                                  "Other Offense", "Percent Minority", 
                                                  "Percent Vote Republican", "Percent Law and Justice")
```


```{r, results='asis', echo=F}
stargazer(upper, header=F, title="Pairwise Correlations", font.size="small", column.sep.width = "0.15pt",
          single.row=TRUE,covariate.labels=c(as.character(paste(c("Variable",1:28)))))
```

\pagebreak


## Table 3 Replication 

The table below presents a replication of Table 3 in Harris et. al. 2011 with data using only felony offenses. A few differences are of note: "SRA score" and "offender score", or comparable measures, are not redily available in the MCAO data. We decided to preliminarily use the number of prior offenses as a proxy for "offender score". Additionally, we use a more expansive offense type categorization for our Native American extentions, including  alcohol-related in addition to violent and drug indicators. We run each of the models on a subset of cases consisting of all felony cases, as to make a closer comparison to the superior court data in Washington.

```{r, include=F}
m3f <- lmer(total_ff_log ~ priors+log(age)+male+black+hispanic+asian+nativeam+other.race+trial_flag+
              drug+alcohol.dui+hunt.fish+other.offense+
              min_perc+repubvote+lecperc+(1|filed_county), REML = F, data = felony)

class(m3f) <- "lmerMod" 
```


```{r, echo=F, results='asis'}
stargazer(m3f, style = "asr", type = "latex", title = "Harris et. al. ASR Table 3 Replication  - MCAO Felony Data",
          covariate.labels = c("Priors", "Age", 
                               "Male",  "Black", "Hispanic", "Asian", "Native Am.", "Other Race", "Trial", 
                               "Alcohol.DUI", "Drug", "Hunt.Fish", "Other Offense", "Percent Minority", 
                               "Percent Vote Republican", "Percent Law and Justice"),
          dep.var.labels =  "Total Fine and Fee Order (log)", model.numbers = F,
          add.lines = list(c("L1 Variance", round(attr(VarCorr(m3f), "sc")^2,2)),
                                      c("L2 Variance", round(as.numeric(VarCorr(m3f)[["filed_county"]]),2))),
          omit.stat = c("aic", "bic", "ll"), header=F)
```

```{r, eval=F, include=F}
summary(m3f)
as.data.frame(ranef(m3f)) %>% select(grp, condval) %>% mutate(exp(condval)) #ui in dollar values
exp(m3f@beta[1]) #baseline
(exp(m3f@beta[-1])-1)*100 #percentage change of b

#diagnostics

#model form - fitted vs residual
plot(m3f)

#linearity - do for each variable
#ggplot(data.frame(x1=monsanc.short$,pearson=residuals(m3,type="pearson")),
#     aes(x=x1,y=pearson))

#normality - qqplot
qqnorm(resid(m3f)) %>% qqline(resid(m3f))

#leverage
influenceIndexPlot(m3f)

#ICC
r1Var <- as.numeric(VarCorr(m3f)[["filed_county"]])
residVar <- attr(VarCorr(m3f), "sc")^2
r1Var
residVar
r1Var / (r1Var + residVar) #only 11.5% of variation is due to county
```

```{r, include=F}
#table 4 - Black Interactions
m41f <- lmer(total_ff_log ~ priors+log(age)+male+black+trial_flag+violent+
              black_perc+repubvote+lecperc+(1|filed_county), REML = F, data = felony)
class(m41f) <- "lmerMod" 

m42f <- lmer(total_ff_log ~ priors+log(age)+male+black+trial_flag+violent+
               black_perc+repubvote+lecperc+black_perc:black+(1|filed_county), 
             REML = F, data = felony)
class(m42f) <- "lmerMod" 

m43f <- lmer(total_ff_log ~ priors+log(age)+male+black+trial_flag+violent+
               black_perc+repubvote+lecperc+violent:black+(1|filed_county), 
             REML = F, data = felony)
class(m43f) <- "lmerMod" 

m44f <- lmer(total_ff_log ~ priors+log(age)+male+black+trial_flag+violent+
               black_perc+repubvote+lecperc+violent:black_perc+(1|filed_county), 
             REML = F, data = felony)
class(m44f) <- "lmerMod" 

```

```{r, echo=F, results='asis'}
stargazer(m41f, m42f, m43f, m44f,
          style = "asr", type = "latex", title = "Harris et. al. ASR Table 4 Replication  - MCAO Felony Data",
          covariate.labels = c("Priors", "Age", 
                               "Male",  "Black",  "Trial", 
                               "Violent", "Percent Black", 
                               "Percent Vote Republican", "Percent Law and Justice",
                               "Black*Percent Black", "Black*Violent", "Violent*Percent Black"),
          dep.var.labels =  "Total Fine and Fee Order (log)", model.numbers = T,
          add.lines = list(c("L1 Variance", round(attr(VarCorr(m41f), "sc")^2,2),
                             round(attr(VarCorr(m42f), "sc")^2,2), round(attr(VarCorr(m43f), "sc")^2,2),
                             round(attr(VarCorr(m44f), "sc")^2,2)),
                                      c("L2 Variance", 
                                        round(as.numeric(VarCorr(m41f)[["filed_county"]]),2),
                                        round(as.numeric(VarCorr(m42f)[["filed_county"]]),2),
                                        round(as.numeric(VarCorr(m43f)[["filed_county"]]),2),
                                        round(as.numeric(VarCorr(m44f)[["filed_county"]]),2))),
          omit.stat = c("aic", "bic", "ll"), header=F)
```

```{r, include=F}
#table 5 - Hispanic Interactions

m51f <- lmer(total_ff_log ~ priors+log(age)+male+hispanic+trial_flag+drug+
               hisp_perc+repubvote+lecperc+(1|filed_county), REML = F, data = felony)
class(m51f) <- "lmerMod" 


m52f <- lmer(total_ff_log ~ priors+log(age)+male+hispanic+trial_flag+drug+
               hisp_perc+repubvote+lecperc+hisp_perc:hispanic+(1|filed_county), 
             REML = F, data = felony)
class(m52f) <- "lmerMod" 

m53f <- lmer(total_ff_log ~ priors+log(age)+male+hispanic+trial_flag+drug+
               hisp_perc+repubvote+lecperc+drug:hispanic+(1|filed_county), 
             REML = F, data = felony)
class(m53f) <- "lmerMod" 

m54f <- lmer(total_ff_log ~ priors+log(age)+male+hispanic+trial_flag+drug+
               hisp_perc+repubvote+lecperc+drug:hisp_perc+(1|filed_county), 
             REML = F, data = felony)
class(m54f) <- "lmerMod" 

```

```{r, echo=F, results='asis'}
stargazer(m51f, m52f, m53f, m54f,
          style = "asr", type = "latex", title = "Harris et. al. ASR Table 5 Replication  - MCAO Felony Data",
          covariate.labels = c("Priors", "Age", 
                               "Male",  "Hispanic",  "Trial", 
                               "Drug", "Percent Hispanic", 
                               "Percent Vote Republican", "Percent Law and Justice",
                               "Hispanic*Percent Hispanic", "Hispanic*Drug", "Drug*Percent Hispanic"),
          dep.var.labels =  "Total Fine and Fee Order (log)", model.numbers = T,
          add.lines = list(c("L1 Variance", round(attr(VarCorr(m51f), "sc")^2,2),
                             round(attr(VarCorr(m52f), "sc")^2,2), round(attr(VarCorr(m53f), "sc")^2,2),
                             round(attr(VarCorr(m54f), "sc")^2,2)),
                                      c("L2 Variance", 
                                        round(as.numeric(VarCorr(m51f)[["filed_county"]]),2),
                                        round(as.numeric(VarCorr(m52f)[["filed_county"]]),2),
                                        round(as.numeric(VarCorr(m53f)[["filed_county"]]),2),
                                        round(as.numeric(VarCorr(m54f)[["filed_county"]]),2))),
          omit.stat = c("aic", "bic", "ll"), header=F)
```

```{r, include=F}
#table "6" - Native American Interactions 
m61f <- lmer(total_ff_log ~ priors+log(age)+male+nativeam+trial_flag+alcohol.dui+
               na_perc+repubvote+lecperc+(1|filed_county), REML = F, data = felony)
class(m61f) <- "lmerMod" 


m62f <- lmer(total_ff_log ~ log(age)+male+nativeam+trial_flag+alcohol.dui+
               na_perc+repubvote+lecperc+na_perc:nativeam+(1|filed_county), 
             REML = F, data = felony)
class(m62f) <- "lmerMod" 

m63f <- lmer(total_ff_log ~ log(age)+male+nativeam+trial_flag+alcohol.dui+
               na_perc+repubvote+lecperc+alcohol.dui:nativeam+hunt.fish:nativeam+(1|filed_county), 
             REML = F, data = felony)
class(m63f) <- "lmerMod" 

m64f <- lmer(total_ff_log ~ log(age)+male+nativeam+trial_flag+alcohol.dui+
               na_perc+repubvote+lecperc+alcohol.dui:na_perc+(1|filed_county), 
             REML = F, data = felony)
class(m64f) <- "lmerMod" 

```

```{r, echo=F, results='asis'}
stargazer(m61f, m62f, m63f, m64f,
          style = "asr", type = "latex", title = "Native American Interactions - MCAO Felony Data",
          covariate.labels = c("Priors", "Age", 
                               "Male",  "Native Am.",  "Trial", 
                               "Alcohol.DUI", "Percent NA", 
                               "Percent Vote Republican", "Percent Law and Justice",
                               "Native Am.*Percent NA", "Native Am.*Alcohol.DUI",
                               "Alcohol.DUI*Percent NA"),
          dep.var.labels =  "Total Fine and Fee Order (log)", model.numbers = T,
          add.lines = list(c("L1 Variance", round(attr(VarCorr(m61f), "sc")^2,2),
                             round(attr(VarCorr(m62f), "sc")^2,2), round(attr(VarCorr(m63f), "sc")^2,2),
                             round(attr(VarCorr(m64f), "sc")^2,2)),
                                      c("L2 Variance", 
                                        round(as.numeric(VarCorr(m61f)[["filed_county"]]),2),
                                        round(as.numeric(VarCorr(m62f)[["filed_county"]]),2),
                                        round(as.numeric(VarCorr(m63f)[["filed_county"]]),2),
                                        round(as.numeric(VarCorr(m64f)[["filed_county"]]),2))),
          omit.stat = c("aic", "bic", "ll"), header=F)
```

