---
title: "Race Paper Final"
author: "Ryan Larson - UMN"
date: "`r format(Sys.time(), '%m/%d/%Y')`"
output:
  pdf_document: default
always_allow_html: true
fontsize: 11pt
editor_options:
  chunk_output_type: console
header-includes: 
  - \usepackage{dcolumn}
  - \pagenumbering{gobble}
---

# Package Preamble

```{r, warning=FALSE, message=FALSE}
library(readr)
library(readxl)
library(tidyr)
library(stringr)
library(dplyr)
library(ggplot2)
library(sem)
library(AER)
library(stargazer)
library(tvthemes)
library(kableExtra)
library(dotwhisker)

#felony only runs
#Figure 2: identify dotted line as white
#first stage in appendix
```

# Data Munging

```{r, warning=F, include=F, message=F, eval=FALSE}
#load data - constructed from MCAO .txt files
monsanc <- read_csv("Data/monsanc.csv") 

#recoding and filtering
monsanc.short <- monsanc %>%
    filter(charge_degree!="Petty Misdemeanor") %>% #no petty misdemeanors
  mutate(total_ff = total_ff*adj, #adjust dollars to Jan. 2018
         total_ff_log = log((total_ff+1)), 
         total_order = total_order*adj,
         total_order_log = log((total_order+1)), 
         conf_days_log = log((conf_days+1)),
         conf_minus_stayed = ifelse(conf_minus_stayed < 0, NA, conf_minus_stayed),
         conf_minus_stayed_log = log((conf_minus_stayed+1)), 
         conf_minus_stayed_ts = ifelse(conf_minus_stayed-credit_days <0, 0, conf_minus_stayed-credit_days),
         conf_minus_stayed_ts_log = log((conf_minus_stayed_ts+1)),
         prob_days_log = log((prob_days+1)),
         perc_credit = ifelse(is.nan(perc_credit), 0, perc_credit),
         perc_credit = ifelse(is.infinite(perc_credit), NA, perc_credit),
         perc_stayed = (conf_days-conf_minus_stayed)/conf_days*100,
         perc_stayed = ifelse(is.nan(perc_stayed), 0, perc_stayed)) %>%
  #filter(file_year >= 2010 & file_year <= 2015 & sentence_year >=2010) %>%
  filter(age >= 15) %>%  #filtering cases with likely error ages 
  filter(prob_days <= 20*365) %>% #filtering cases with likely error probation (above 65 years)
  filter(credit_days < 129564) %>% #removing removing likely error credit case
  mutate(charge_degree = relevel(factor(charge_degree, ordered = F), ref = "Misdemeanor"),
         charge_offense = relevel(factor(charge_offense, ordered = F), ref = "violent"),
         filed_district = relevel(factor(
         filed_district, ordered = F), ref = "04"), 
         felony = ifelse(charge_degree=="Felony", 1, 0), 
         gross.mis = ifelse(charge_degree=="Gross Misdemeanor", 1, 0),
         mis = ifelse(charge_degree=="Misdemeanor", 1, 0),
         petty.mis = ifelse(charge_degree=="Petty Misdemeanor", 1, 0),
         violent = ifelse(charge_offense=="violent", 1, 0),
         alcohol.dui = ifelse(charge_offense=="alcohol/dui", 1, 0),
         drug = ifelse(charge_offense=="drug", 1, 0),
         hunt.fish = ifelse(charge_offense=="hunt/fish", 1, 0),
         other.offense = ifelse(charge_offense=="other", 1, 0),
         male = ifelse(gender_impute=="M", 1,0),
         perc_credit = ifelse(perc_credit > 100, 100, perc_credit),
         trial_flag = ifelse(trial_flag==TRUE, 1,0),
         pubdef = ifelse(pubdef=="yes", 1,0),
         felony_flag = ifelse(felony_flag==TRUE,1,0),
         gm_flag = ifelse(gm_flag==TRUE,1,0),
         violent_flag = ifelse(violent_flag==TRUE,1,0),
         drug_flag = ifelse(drug_flag==TRUE,1,0),
         alcohol_flag = ifelse(alcohol_flag==TRUE,1,0)) %>%
  fastDummies::dummy_cols(select_columns = c("filed_district", "sentence_year"))

#% Missing on defendant-level demographics (p. 19 of draft)
prop.table(table(is.na(monsanc.short$race_impute)))
prop.table(table(is.na(monsanc.short$gender_impute)))

#defining focal race factor
monsanc.short <- monsanc.short %>%
  mutate(race_impute = case_when(
    race_impute=="white"~"white",
    race_impute=="black"~"black",
    race_impute=="hispanic"~"hispanic",
    race_impute=="asian"~"asian",
    race_impute=="nat. am."~"nat. am.",
    race_impute=="other"~"other",
    is.na(race_impute)~"missing"
  )) %>%
  mutate(race_impute = relevel(factor(race_impute, ordered = F), 
                               ref = "white"),
         race.miss = ifelse(race_impute=="missing", 1, 0),
          white = ifelse(race_impute=="white", 1, 0),
         asian = ifelse(race_impute=="asian", 1, 0),
         black = ifelse(race_impute=="black", 1, 0),
         hispanic = ifelse(race_impute=="hispanic", 1, 0),
         nativeam = ifelse(race_impute=="nat. am.", 1, 0),
         other.race = ifelse(race_impute=="other", 1, 0))

#joining and munging capacity data
capacity <- read_xlsx("Data/capacity.xlsx") %>%
  filter(year >= 2004 & year <= 2018 & state=="MN") %>% 
  dplyr::select(county_name, year, total_jail_pop, jail_rated_capacity) %>% 
  mutate(cap_ratio = total_jail_pop/jail_rated_capacity) %>%
  mutate(cap_ratio = ifelse(is.nan(cap_ratio), 1, cap_ratio)) %>%
  group_by(county_name) %>%
  mutate(cap_ratio_lead = dplyr::lead(cap_ratio, 1, order_by = year)) %>%
  ungroup()

regional_jails <- read_xlsx("Data/capacity.xlsx") %>%
  filter(year >= 2004 & year <= 2018 & state=="MN") %>%
  filter(regional_jail_flag==1) %>%
  pull(county_name) %>%
  unique()
  

monsanc.short <- monsanc.short %>% 
  left_join(capacity, by = c("filed_county"="county_name", "sentence_year"="year")) %>%
  mutate(regional_jail = ifelse(filed_county %in% regional_jails, 1, 0))

#fix pubdef variable
table(monsanc.short$pubdef, monsanc.short$attorney, useNA = "always") #no lawyer stated as missing in pubdef

monsanc.short <- monsanc.short %>%
  mutate(pubdef = ifelse(is.na(pubdef), 0, pubdef),
         perc_credit = ifelse(is.na(perc_credit), 0, perc_credit))

#listwise deletion of missing data
#restrict cases with felony and gross misdemeanor highest charges
monsanc.short <- monsanc.short  %>%
  drop_na(conf_minus_stayed_ts, total_ff, prob_days,
          age, trial_flag,
        priors, pubdef, perc_credit,
        perc_stayed, 
        felony_flag, gm_flag, 
        violent_flag, drug_flag, alcohol_flag,
        filed_district, sentence_year,
        cap_ratio) %>%
  filter(charge_degree=="Felony"|
           charge_degree=="Gross Misdemeanor")

#listwise deletion demographic missing
monsanc.short <- monsanc.short  %>%
  drop_na(race_impute, gender_impute) 

write_csv(monsanc.short, "Data/monsanc.short.csv")
```


# Analysis with Gross Misdemeanor and Felony Highest Charge Cases

```{r}
monsanc.short <- read_csv("Data/monsanc.short.csv")
```

## Table 1: Descriptive Statistics

```{r, results='asis'}
monsanc.ds <- as.data.frame(monsanc.short[,c("conf_minus_stayed_ts",
                                         "total_ff", 
                                  "prob_days",  
                                 "white", "black", "hispanic",
                                 "asian", "nativeam",  
                                 "other.race", "race.miss",
                                 "male", "age", "trial_flag",
                                 "priors", "pubdef", "perc_credit",
                                 "perc_stayed",
                                 "felony_flag", "gm_flag",
                                 "violent_flag", "drug_flag","alcohol_flag",
                                 "filed_district_01", "filed_district_02",  
                                 "filed_district_03", 
                                 "filed_district_04", "filed_district_05",  
                                 "filed_district_06", "filed_district_07",  
                                 "filed_district_08", "filed_district_09", 
                                 "filed_district_10", 
                                 "sentence_year_2004","sentence_year_2005",
                                 "sentence_year_2006","sentence_year_2007",
                                 "sentence_year_2008","sentence_year_2009",
                                 "sentence_year_2010","sentence_year_2011",
                                 "sentence_year_2012",
                                 "sentence_year_2013", "sentence_year_2014",
                                 "sentence_year_2015", "sentence_year_2016",
                                 "sentence_year_2017",
                                 "cap_ratio")])

stargazer(monsanc.ds,
          covariate.labels = c("Incarceration Days",
                               "Total Fine/Fee Order", 
                               "Probation Days",  
                                 "White", "Black", "Hispanic",
                                 "Asian", "Native American",  
                                 "Other Race", "Missing Race", 
                                 "Male", "Age", "Trial",
                                 "Priors", "Public Defender", 
                                 "Percent Credit", "Percent Stayed",
                                 "Felony", "Gross Misdemeanor",
                                 "Violent", "Drug","Alcohol/DUI",
                               "Judicial District 1", "Judicial District 2",
                               "Judicial District 3", "Judicial District 4",
                               "Judicial District 5", "Judicial District 6",
                               "Judicial District 7", "Judicial District 8",
                               "Judicial District 9", 
                               "Judicial District 10",
                               "Year - 2004", "Year - 2005",
                               "Year - 2006", "Year - 2007",
                               "Year - 2008", "Year - 2009",
                               "Year - 2010", 
                               "Year - 2011", "Year - 2012",
                               "Year - 2013", "Year - 2014", "Year - 2015",
                               "Year - 2016", "Year - 2017",
                               "County-Level Capacity Ratio"),
          type="latex", 
          style="asr", 
          title="Descriptive Statistics for Variables SCAO and VERA Jail Data",
          summary=T, 
          median=T,  
          header = F)
```

## Figure 1: Punishment Amounts by Race and Type

```{r, message=FALSE}
#faceted bar graph
fig1 <- monsanc.short %>%
  select(race_impute, total_ff,
         conf_minus_stayed_ts, prob_days) %>%
 mutate(Race = case_when(
    race_impute=="asian"~"Asian",
    race_impute=="black"~"Black",
    race_impute=="hispanic"~"Hispanic",
    race_impute=="nat. am."~"Nat. Am.",
    race_impute=="other"~"Other",
    race_impute=="white"~"White"
  )) %>%
  filter(!is.na(Race)) %>%
  select(-race_impute) %>%
  group_by(Race) %>%
  summarize(LFO = mean(total_ff, na.rm = T),
            Incarceration = mean(conf_minus_stayed_ts, na.rm = T),
            Probation = mean(prob_days, na.rm = T)) %>%
  mutate(Race = factor(Race, 
                       levels = c("White", "Black",
                                  "Hispanic", "Asian", 
                                  "Nat. Am.", "Other"))) %>%
  pivot_longer(cols = c("LFO", "Incarceration", "Probation"),
               names_to = "punishment",
               values_to = "amount") %>%
  ggplot()+
  geom_bar(aes(x=Race, y=amount, fill = Race),
           color = "black",
           stat="identity", 
           position = position_dodge2())+
  geom_text(aes(x=Race, y=amount+25,
              label = round(amount,0)), 
          position = position_dodge2(width = 1))+
  facet_wrap(~punishment)+
  labs(title = "Figure 1: Punishment Amounts by Defendant Race",
       subtitle = "MN SCAO 2011-2015",
       y = "Amount (Days/USD/Days)")+
  tvthemes::scale_fill_westeros(palette = "Stark")+
  theme_classic()+
  theme(legend.position = "none"#,
        #text=element_text(family="Times New Roman")
        )

fig1
```

## Table 2: Multivariate Regression of Punishment

$$Y_{ik}=\alpha_k + \sum \pi_{jk}Race_{ij}+\sum \beta_{jk}X_{ij}+\theta_{dk}+\lambda_{tk}+\epsilon_{ik}$$

```{r}
#Multivariate Regression - when all predictors are identical across models
  #equivalent to separate OLS, but more efficient and takes into account error covariance

#multivariate regression
mv <- lm(cbind(log(conf_minus_stayed_ts_log+1),
               log(prob_days+1),
               log(total_ff+1))~
                 black+hispanic+asian+nativeam+other.race+
                 race.miss+
                 male+log(age)+
                  priors+pubdef+perc_credit+
           perc_stayed+trial_flag+
                 felony_flag+gm_flag+
                 violent_flag+drug_flag+alcohol_flag+
                 filed_district+
                 as.factor(sentence_year), 
         data = monsanc.short)

summary(mv)

conf_model <- lm(log(conf_minus_stayed_ts+1)~
                 black+hispanic+asian+nativeam+other.race+
                 race.miss+
                 male+log(age)+
                  priors+pubdef+perc_credit+
           perc_stayed+trial_flag+
                 felony_flag+gm_flag+
                 violent_flag+drug_flag+alcohol_flag+
                 filed_district+
                 as.factor(sentence_year), 
         data = monsanc.short)

prob_model <- lm(log(prob_days+1)~
                  black+hispanic+asian+nativeam+other.race+
                 race.miss+
                 male+log(age)+
                  priors+pubdef+perc_credit+
           perc_stayed+trial_flag+
                 felony_flag+gm_flag+
                 violent_flag+drug_flag+alcohol_flag+
                 filed_district+
                 as.factor(sentence_year), 
         data = monsanc.short)

lfo_model <- lm(log(total_ff+1)~
                  black+hispanic+asian+nativeam+other.race+
                 race.miss+
                 male+log(age)+
                  priors+pubdef+perc_credit+
           perc_stayed+trial_flag+
                 felony_flag+gm_flag+
                 violent_flag+drug_flag+alcohol_flag+
                 filed_district+
                 as.factor(sentence_year), 
         data = monsanc.short)
```


```{r, results='asis'}
#stargazer regression table

stargazer(conf_model, lfo_model, prob_model,
           type = "latex", 
          title = "Multivariate Model of Punishment, Minnesota 2004-2017",
          covariate.labels = c("Black", "Hispanic", "Asian",
                               "Native American", "Other Race",
                               "Missing Race", "Male", "log(Age)",
                               "Prior Convictions", "Public Defender", 
                               "Percent Credit", "Percent Stayed",
                               "Trial", 
                               "Felony", "Gross Misdemeanor", 
                               "Violent", "Drug", "Alcohol/DUI"),
          model.numbers = FALSE,
          header = FALSE,
          dep.var.caption  = "Punishment Outcome",
          dep.var.labels = c("log(Incarceration)", 
                            "log(LFO)",
                            "log(Probation)"),
          column.labels = c("Coef(SE)", "Coef(SE)", "Coef(SE)"),
          single.row = TRUE,
          font.size="footnotesize", 
          no.space = T, 
          column.sep.width = "1pt",
          align = TRUE,
          omit.stat = c("ser"),
          omit = c("filed_district", "sentence_year"),
          star.cutoffs = c(.05, .01, .001), 
          star.char = c("*","**","***"),
          add.lines = list(c("District FE", "Yes", "Yes", "Yes"),
                           c("Sentence Year FE", "Yes", "Yes", "Yes")))
```

## Table 3: Instrumental Variable Models of Punishment

$$I_i = \alpha + \phi_{1}CR_{d} +\sum \beta_{j}X_{ij}+\theta_{d}+\lambda_{t}+\epsilon_{i} $$

$$Y_{ik} = \alpha + \phi_{2k} CR_d+\sum \pi_{jk}Race_{ij}+\sum \beta_{jk}X_{ij}+\theta_{dk}+\lambda_{tk}+\epsilon_{ik}$$

$$\delta_k = \frac{\phi_{2k}}{\phi_{1}}$$

```{r}
#instrumental Variables Regression 

#LFO
iv.ff <- ivreg(total_ff_log~conf_minus_stayed_ts_log+
                 black+hispanic+asian+nativeam+other.race+
                 race.miss+
                 male+log(age)+
                  priors+pubdef+perc_credit+
                 perc_stayed+
                 trial_flag+
                 felony_flag+gm_flag+
                 violent_flag+drug_flag+alcohol_flag+
                 filed_district+
                 as.factor(sentence_year)|
              .-conf_minus_stayed_ts_log+cap_ratio, 
           data = monsanc.short)

summary(iv.ff, diagnostics=T)

#Probation
iv.prob <- ivreg(prob_days_log~conf_minus_stayed_ts_log+
                 black+hispanic+asian+nativeam+other.race+
                 race.miss+
                 male+log(age)+
                 priors+pubdef+perc_credit+
                 perc_stayed+
                 trial_flag+
                 felony_flag+gm_flag+
                 violent_flag+drug_flag+alcohol_flag+
                 filed_district+
                 as.factor(sentence_year)|
              .-conf_minus_stayed_ts_log+cap_ratio, 
           data = monsanc.short)

summary(iv.prob, diagnostics=T)

#exogeneity checks

ppcor::pcor.test(y = monsanc.short$cap_ratio_lead[!is.na(monsanc.short$cap_ratio_lead)], 
                 x = monsanc.short$total_ff_log[!is.na(monsanc.short$cap_ratio_lead)],
                 z = monsanc.short$cap_ratio[!is.na(monsanc.short$cap_ratio_lead)],
                 method = "pearson")

ppcor::pcor.test(y = monsanc.short$cap_ratio_lead[!is.na(monsanc.short$cap_ratio_lead)], 
                 x = monsanc.short$prob_days_log[!is.na(monsanc.short$cap_ratio_lead)],
                 z = monsanc.short$cap_ratio[!is.na(monsanc.short$cap_ratio_lead)],
                 method = "pearson")
```

```{r, results='asis'}
stargazer(iv.ff, iv.prob,
          type = "latex", 
          title = "IV 2SLS Models of Punishment, Minnesota 2004-2017",
          covariate.labels = c("log(Incarceration)",
                               "Black", "Hispanic", "Asian",
                               "Native American", "Other Race",
                               "Missing Race", "Male", "log(Age)",
                               "Prior Convictions", "Public Defender", 
                               "Percent Credit", "Percent Stayed",
                               "Trial", 
                               "Felony", "Gross Misdemeanor", 
                               "Violent", "Drug", "Alcohol/DUI"),
          model.numbers = FALSE,
          header = FALSE,
         dep.var.caption  = "Punishment Outcome",
          dep.var.labels = c("log(LFO)",
                            "log(Probation)"),
          column.labels = c("Coef(SE)", "Coef(SE)"),
          single.row = TRUE,
          font.size="footnotesize", 
          no.space = T, 
          column.sep.width = "1pt",
          align = TRUE,
          omit = c("filed_district", "sentence_year"),
          omit.stat = c("adj.rsq", "rsq", "ser"),
          star.cutoffs = c(.05, .01, .001), 
          star.char = c("*","**","***"),
          add.lines = list(c("District FE", "Yes", "Yes", "Yes"),
                           c("Sentence Year FE", "Yes", "Yes", "Yes"),
                           c("IV F(Incar.)", "10.58^{**}", "10.58^{**}"),
                           c("IV Wu-Hausman", "136.08^{***}", "57.62^{***}")),
          notes = "IV: County-Level Jail Capacity Ratio")
```

## Figure 2: Coefficient Plot

```{r}
conf_coef_mv <- broom::tidy(conf_model) %>%
  filter(str_detect(term, pattern = "black|hispanic|asian|nativeam|other.race")) %>%
  relabel_predictors(c(
    black = "Black",
    hispanic = "Hispanic",
    asian = "Asian",
    nativeam = "Native American",
    other.race = "Other Race"
  )) %>%
  mutate(model = "MV",
         punishment = "Incarceration")

prob_coef_mv <- broom::tidy(prob_model) %>%
  filter(str_detect(term, pattern = "black|hispanic|asian|nativeam|other.race")) %>%
  relabel_predictors(c(
    black = "Black",
    hispanic = "Hispanic",
    asian = "Asian",
    nativeam = "Native American",
    other.race = "Other Race"
  )) %>%
  mutate(model = "MV",
         punishment = "Probation")

lfo_coef_mv <- broom::tidy(lfo_model) %>%
  filter(str_detect(term, pattern = "black|hispanic|asian|nativeam|other.race")) %>%
  relabel_predictors(c(
    black = "Black",
    hispanic = "Hispanic",
    asian = "Asian",
    nativeam = "Native American",
    other.race = "Other Race"
  )) %>%
  mutate(model = "MV",
         punishment = "LFO")

prob_coef_iv <- broom::tidy(iv.prob) %>%
  filter(str_detect(term, pattern = "black|hispanic|asian|nativeam|other.race")) %>%
  relabel_predictors(c(
    black = "Black",
    hispanic = "Hispanic",
    asian = "Asian",
    nativeam = "Native American",
    other.race = "Other Race"
  )) %>%
  mutate(model = "IV",
         punishment = "Probation")

lfo_coef_iv <- broom::tidy(iv.ff) %>%
  filter(str_detect(term, pattern = "black|hispanic|asian|nativeam|other.race")) %>%
  relabel_predictors(c(
    black = "Black",
    hispanic = "Hispanic",
    asian = "Asian",
    nativeam = "Native American",
    other.race = "Other Race"
  )) %>%
  mutate(model = "IV",
         punishment = "LFO")

mv_coef <- rbind(conf_coef_mv, 
                 prob_coef_mv,
                 lfo_coef_mv,
                 prob_coef_iv, 
                 lfo_coef_iv) %>%
  mutate(model = factor(model, levels = c("MV", "IV")))

dwplot(mv_coef,
       vline = geom_vline(
        xintercept = 0,
        colour = "grey60",
        linetype = 2),
        dot_args = list(aes(shape = model)),
    whisker_args = list(aes(linetype = model))) +
  theme_classic()+
  #theme(text=element_text(family="Times New Roman"))+
facet_wrap(~punishment)+
  labs(x = "Coefficient Estimate",
       y = "",
       title = "Figure 2: Coefficient Plots for MV and IV Punishment Models",
       subtitle = "MV = Multivariate, IV = Instrumental Variable",
       caption = "Dotted line represents the referent group - White defendants.") +
  guides(shape = guide_legend("Model"), 
        colour = guide_legend("Model"))+
    scale_colour_manual(
        values = c("black", "orange"),
        name = "Model",
        breaks = c("MV", "IV"),
        labels = c("MV", "IV")) +
    scale_shape_manual(
      values = c(1,2),
        name = "Model",
        breaks = c("MV", "IV"),
        labels = c("MV", "IV"))
```

# Appendix

## Robustness Models without Regional Jails

### Table A1: Multivariate Model of Punishment w/o Regional Jails

```{r}
conf_model_rj <- lm(log(conf_minus_stayed_ts+1)~
                 black+hispanic+asian+nativeam+other.race+
                 race.miss+
                 male+log(age)+
                  priors+pubdef+perc_credit+
           perc_stayed+trial_flag+
                 felony_flag+gm_flag+
                 violent_flag+drug_flag+alcohol_flag+
                 filed_district+
                 as.factor(sentence_year), 
         data = monsanc.short[monsanc.short$regional_jail==0,])

prob_model_rj <- lm(log(prob_days+1)~
                  black+hispanic+asian+nativeam+other.race+
                 race.miss+
                 male+log(age)+
                  priors+pubdef+perc_credit+
           perc_stayed+trial_flag+
                 felony_flag+gm_flag+
                 violent_flag+drug_flag+alcohol_flag+
                 filed_district+
                 as.factor(sentence_year), 
          data = monsanc.short[monsanc.short$regional_jail==0,])

lfo_model_rj <- lm(log(total_ff+1)~
                  black+hispanic+asian+nativeam+other.race+
                 race.miss+
                 male+log(age)+
                  priors+pubdef+perc_credit+
           perc_stayed+trial_flag+
                 felony_flag+gm_flag+
                 violent_flag+drug_flag+alcohol_flag+
                 filed_district+
                 as.factor(sentence_year), 
          data = monsanc.short[monsanc.short$regional_jail==0,])
```


```{r, results='asis'}
#stargazer regression table

stargazer(conf_model_rj, lfo_model_rj, prob_model_rj,
           type = "latex", 
          title = "Multivariate Model of Punishment w/o Regional Jail Counties, Minnesota 2004-2017",
          covariate.labels = c("Black", "Hispanic", "Asian",
                               "Native American", "Other Race",
                               "Missing Race", "Male", "log(Age)",
                               "Prior Convictions", "Public Defender", 
                               "Percent Credit", "Percent Stayed",
                               "Trial", 
                               "Felony", "Gross Misdemeanor", 
                               "Violent", "Drug", "Alcohol/DUI"),
          model.numbers = FALSE,
          header = FALSE,
          dep.var.caption  = "Punishment Outcome",
          dep.var.labels = c("log(Incarceration)", 
                            "log(LFO)",
                            "log(Probation)"),
          column.labels = c("Coef(SE)", "Coef(SE)", "Coef(SE)"),
          single.row = TRUE,
          font.size="footnotesize", 
          no.space = T, 
          column.sep.width = "1pt",
          align = TRUE,
          omit.stat = c("ser"),
          omit = c("filed_district", "sentence_year"),
          star.cutoffs = c(.05, .01, .001), 
          star.char = c("*","**","***"),
          add.lines = list(c("District FE", "Yes", "Yes", "Yes"),
                           c("Sentence Year FE", "Yes", "Yes", "Yes")))
```

### Table A2: Instrumental Variable Model of Punishment w/o Regional Jails

```{r}
#instrumental Variables Regression 

#LFO
iv.ff.rj <- ivreg(total_ff_log~conf_minus_stayed_ts_log+
                 black+hispanic+asian+nativeam+other.race+
                 race.miss+
                 male+log(age)+
                  priors+pubdef+perc_credit+
                 perc_stayed+
                 trial_flag+
                 felony_flag+gm_flag+
                 violent_flag+drug_flag+alcohol_flag+
                 filed_district+
                 as.factor(sentence_year)|
              .-conf_minus_stayed_ts_log+cap_ratio, 
           data = monsanc.short[monsanc.short$regional_jail==0,])

summary(iv.ff.rj, diagnostics=T)

#Probation
iv.prob.rj <- ivreg(prob_days_log~conf_minus_stayed_ts_log+
                 black+hispanic+asian+nativeam+other.race+
                 race.miss+
                 male+log(age)+
                 priors+pubdef+perc_credit+
                 perc_stayed+
                 trial_flag+
                 felony_flag+gm_flag+
                 violent_flag+drug_flag+alcohol_flag+
                 filed_district+
                 as.factor(sentence_year)|
              .-conf_minus_stayed_ts_log+cap_ratio, 
           data = monsanc.short[monsanc.short$regional_jail==0,])

summary(iv.prob.rj, diagnostics=T)
```

```{r, results='asis'}
stargazer(iv.ff.rj, iv.prob.rj,
          type = "latex", 
          title = "IV 2SLS Models of Punishment w/o Regional Jail Counties, Minnesota 2004-2017",
          covariate.labels = c("log(Incarceration)",
                               "Black", "Hispanic", "Asian",
                               "Native American", "Other Race",
                               "Missing Race", "Male", "log(Age)",
                               "Prior Convictions", "Public Defender", 
                               "Percent Credit", "Percent Stayed",
                               "Trial", 
                               "Felony", "Gross Misdemeanor", 
                               "Violent", "Drug", "Alcohol/DUI"),
          model.numbers = FALSE,
          header = FALSE,
         dep.var.caption  = "Punishment Outcome",
          dep.var.labels = c("log(LFO)",
                            "log(Probation)"),
          column.labels = c("Coef(SE)", "Coef(SE)"),
          single.row = TRUE,
          font.size="footnotesize", 
          no.space = T, 
          column.sep.width = "1pt",
          align = TRUE,
          omit = c("filed_district", "sentence_year"),
          omit.stat = c("adj.rsq", "rsq", "ser"),
          star.cutoffs = c(.05, .01, .001), 
          star.char = c("*","**","***"),
          add.lines = list(c("District FE", "Yes", "Yes", "Yes"),
                           c("Sentence Year FE", "Yes", "Yes", "Yes"),
                           c("IV F(Incar.)", "4.29^{*}", "4.29^{*}"),
                           c("IV Wu-Hausman", "159.444^{***}",
                             "74.99^{***}")),
          notes = "IV: County-Level Jail Capacity Ratio")
```